LCS Tool: A Computational Platform for Lagrangian Coherent Structures

### 2014-12-11 main.tex
Remove & revise FIXMEs
Conflicts:
	main.tex

Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$. \citet{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
\begin{gather*}
Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{\langle r'(s),D_{t_0}^t(r(s)) r'(s)\rangle}{\sqrt{\langle r'(s),C_{t_0}^t(r(s)) r'(s)\rangle\langle r'(s),r'(s)\rangle}}\dif s,\\
D_{t_0}^t = \frac12 [C_{t_0}^t \Omega - \Omega C_{t_0}^t],
\end{gather*}
% FIXME Clarify what is meant by _closed_ material lines
obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along closed material lines parametrized as $r(s)$ with $s \in [0,\sigma]$. More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations. We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \citet{farazmand14:_shearless}).

Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_1(r),
\label{eq:shrink line}
\end{equation}
and attracting LCSs (\emph{stretch lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_2(r).
\label{eq:stretch line}
\end{equation}

Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$. For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.

@@

% FIXME Need to update table to account for lambda range
\begin{table*}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{.95\textwidth}|}
{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
\hline \hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!lambda! & scalar lambda value in \cref{eq:eta}\tabularnewline
\hline \hline
\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
\hline \hline
\lstinline!domain! & array to define flow domain\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define main grid resolution for Cauchy-Green strain tensor\tabularnewline
\hline
\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\tabularnewline
\hline
\lstinline!PSList! & user-defined structure for Poincare section end-points, number of $\lambda$-lines launched from Poincare section, and maximum closed $\lambda$-line length\tabularnewline
\hline
\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default: 5.\tabularnewline
\hline
\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default: $10^{-2}$.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify $\lambda$-line integration parameters\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
\hline
\end{tabular}
\caption{Syntax for LCS Tool elliptic LCS functions.}
\label{t:Elliptic LCS functions}
\end{table*}

The first step is to define the position of Poincare sections in regions where closed $\lambda$-lines are expected based on a visual analysis of the orbit structure of the $\eta_\pm^\lambda$ vector field. The Poincare section is to be oriented such that the first endpoint is close to the centre of the expected Lagrangian vortex, and the second endpoint is outside this vortex. There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \citet{karrasch14:_autom_lagran} for a recently developed algorithm). Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined. A reasonable default value is 100.

@@

% FIXME Citations without pages (karrasch14:_autom_lagran) do not list publication year
\bibliographystyle{elsarticle-num-names} 
\bibliography{main}

>>>>

Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$. \citet{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
\begin{gather*}
Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{\langle r'(s),D_{t_0}^t(r(s)) r'(s)\rangle}{\sqrt{\langle r'(s),C_{t_0}^t(r(s)) r'(s)\rangle\langle r'(s),r'(s)\rangle}}\dif s,\\
D_{t_0}^t = \frac12 [C_{t_0}^t \Omega - \Omega C_{t_0}^t],
\end{gather*}
obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \citet{farazmand14:_shearless}).

Solutions to this variational problem turn out to be curves of the $\xi_1$ or $\xi_2$ eigenvector field.
Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_1(r),
\label{eq:shrink line}
\end{equation}
and attracting LCSs (\emph{stretch lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_2(r).
\label{eq:stretch line}
\end{equation}

Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting curves of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.

@@

% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
\begin{table*}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{.95\textwidth}|}
{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
\hline \hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!lambda! & scalar lambda value in \cref{eq:eta}\tabularnewline
\hline \hline
\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
\hline \hline
\lstinline!domain! & array to define flow domain\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define main grid resolution for Cauchy-Green strain tensor\tabularnewline
\hline
\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\tabularnewline
\hline
\lstinline!PSList! & user-defined structure for Poincare section end-points, number of $\lambda$-lines launched from Poincare section, and maximum closed $\lambda$-line length\tabularnewline
\hline
\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default: 5.\tabularnewline
\hline
\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default: $10^{-2}$.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify $\lambda$-line integration parameters\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
\hline
\end{tabular}
\caption{Syntax for LCS Tool elliptic LCS functions.}
\label{t:Elliptic LCS functions}
\end{table*}

The first step is to define the position of Poincare sections in regions where closed $\lambda$-lines are expected based on a visual analysis of the orbit structure of the $\eta_\pm^\lambda$ vector field.
The Poincare section is to be oriented such that the first endpoint is close to the centre of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
A reasonable default value is 100.

@@

\bibliographystyle{elsarticle-num-names} 
\bibliography{main}
